What's hot and what's not:\\Windowed developer topic analysis

### 2009-06-29 lda-paper.tex
migod's edits

  As development on a software project progresses, developers shift
  their focus between different topics and tasks many times. Managers
  and newcomer developers often seek ways of understanding what tasks
  have recently been worked on and how much effort has gone into each;
  for example, a manager might wonder what unexpected tasks occupied
  their team's attention during a period when they were supposed to
  have been implementing new features.  
%  have been implementing a set of new features.  
%
  Tools such as Latent Dirichlet Allocation (LDA) and Latent Semantic
%  Indexing (LSI) can be used to extract a set of independent topics
  Indexing (LSI) extract a set of independent topics
  from a corpus of commit-log comments.  Previous work in the area has
  created a single set of topics by analyzing comments from the entire
  lifetime of the project.  In this paper, we propose windowing the
  topic analysis to give a more nuanced view of the system's
  evolution. By using a defined time-window of, for example, one
  month, we can track which topics come and go over time, and which
  ones recur.
%
%
%   Tools such as Latent
%   Dirichlet Allocation (LDA) and Latent Semantic Indexing (LSI) can be
%   used to extract a set of independent topics from a corpus of commit
%   log comments.  Previous work in the area has created a single set of
%   topics by analyzing comments from the entire lifetime of the
%   project.  In this paper, we propose windowing the topic analysis to
%   give a more nuanced view of the system's evolution. By using a
%   defined time-window of, for example, one month, we can track which
%   topics come and go over time, and which ones recur.
%  
  We propose
  visualizations of this model that allows us to explore the evolving
  stream of topics of development occurring over time. We demonstrate
  that windowed topic analysis offers advantages over topic analysis
  applied to a project's lifetime because many topics are quite
  local.

@@

Previous work on topic analysis has been applied to the entire history of a 
project~\cite{1321709-short,lukins2008,10.1109/ICPC.2007.13-short}
to produce a single set of topics. In this paper we explore the idea
of time-windowed topic analysis; that is, we pick a period of, say,
one month and perform topic analysis within each window.  In addition
to being able to study just which topics were actually worked on and
when, we can also analyze the historical timeline of topics for
patterns and trends.  For example, we might wish to categorize each
topic as being local to a single time window, spanning multiple
consecutive windows, or recurring periodically.  Alternatively, we
might explore if any topics often recur together, if one particular
topic usually precedes another, if unplanned topics outnumber planned
topics, if particular topics recur near the end of development
iterations, etc.  whereas this paper suggests that most topics are
local.

@@

% We want to see topics that are unique to a
% certain time period, displayed across the time axis.
% Figure \ref{fig:lda} illustrates the kind of visualization we want to
% automatically generate. 
As part of this work, we wish to explore approaches to automatically
generating visualizations of the topic trends and development timelines, so
that managers may be able to more easily grasp the topic-related
activities of the developers.  Figure \ref{fig:lda} gives an example.
Those topics that
recur, or occur over a larger period are plotted continuously. In our
example we have titled each topic with a word drawn from its
word distribution.  
Given a visualization, such as Figure \ref{fig:lda}, topic
analysis can aid in partitioning a project's time-line into periods of
developer focus.
By breaking apart an iteration into sets of topics
and trends (topics that recur), we may be able to recognize the
underlying software development process and maintenance tasks from these
commits.

@@

Our contributions include:
\begin{itemize}
\item We demonstrate that windowed topic analysis most topics are local
  in scope, which would be missed by a global topic analysis.
\item We present a number of visualizations of topics and their trends over time to aid
  communication and analysis of these topics and trends.
\item We provide an exploratory case study using these techniques on
  several open source database systems.
\end{itemize}

@@

Lukins, Kraft and Etzkorn~\cite{lukins2008} use LDA to help bug
localization by querying for documents that are related to a bug's
topic. They used LDA to build a topic model and then searched for
similar documents by querying with sample documents.
%These queries would indicate the proportion of
%topics that were associated with the query and the
%related documents.

@@

In addition to finding the topics, LDA can also associate the
documents (i.e., commit messages) in its corpus to one or more of the
topics it discovers.  For example, a commit of a simple bug fix would
probably contain a word such as \emph{ticket}, and thus heavily associate
with the bug fix topic.  A commit that concerned user documentation of
the GUI would likely be associated with both the documentation and GUI
topics if it shared words from both topics.  And a commit that did not
concern bugs, GUIs, or documentation would likely be related to the other topic.

@@

Once we have our data windowed, we apply our Topic Analysis tool to
each window and extract the top $N$ topics.  We decided to use a
threshold of $20$ topics based on our previous experience with LDA; we
found that using more than $20$ makes the boundaries between topics less
distinct.

@@

\shrinkit
\SubSection{Visualization}
\shrinkit
%       1. [ ] need the visualizer
  Visualization is an integral part of our topics analysis which
  allows us to quickly explore the topics and trends of a project.
  Visualizing the results %is desirable as it 
  allows us to explore the
  data from different points of view to address different questions:
  How are the topics spread over time?  How many independent topics
  per period were there?  What were the repeating trends?  Did the
  trends dominate, or did local topics dominate?  How broken up were
  the trends?  
  To address these multiple questions we will describe and employ multiple visualizations.
%  Since we have a variety of questions to answer we suggest multiple
%  visualizations, which we now describe.

@@

Adding and removing stop words produced different results. Our choice
of stop words could be biased, and could affect the results.
%, and our bias for or against such stop words could affect
%the results.

>>>>

  As development on a software project progresses, developers shift
  their focus between different topics and tasks many times. Managers
  and newcomer developers often seek ways of understanding what tasks
  have recently been worked on and how much effort has gone into each;
  for example, a manager might wonder what unexpected tasks occupied
  their team's attention during a period when they were supposed to
  have been implementing new features.  
%  have been implementing a set of new features.  
%
  Tools such as Latent Dirichlet Allocation (LDA) and Latent Semantic
%  Indexing (LSI) can be used to extract a set of independent topics
  Indexing (LSI) can be used to extract a set of independent topics
  from a corpus of commit-log comments.  Previous work in the area has
  created a single set of topics by analyzing comments from the entire
  lifetime of the project.  In this paper, we propose windowing the
  topic analysis to give a more nuanced view of the system's
  evolution. By using a defined time-window of, for example, one
  month, we can track which topics come and go over time, and which
  ones recur.
%
%
%   Tools such as Latent
%   Dirichlet Allocation (LDA) and Latent Semantic Indexing (LSI) can be
%   used to extract a set of independent topics from a corpus of commit
%   log comments.  Previous work in the area has created a single set of
%   topics by analyzing comments from the entire lifetime of the
%   project.  In this paper, we propose windowing the topic analysis to
%   give a more nuanced view of the system's evolution. By using a
%   defined time-window of, for example, one month, we can track which
%   topics come and go over time, and which ones recur.
%  
  We propose
  visualizations of this model that allows us to explore the evolving
  stream of topics of development occurring over time. We demonstrate
  that windowed topic analysis offers advantages over topic analysis
  applied to a project's lifetime because many topics are quite
  local.

@@

Previous work on topic analysis has been applied to the entire history of a 
project~\cite{1321709-short,lukins2008,10.1109/ICPC.2007.13-short}
to produce a single set of topics. In this paper we explore the idea
of time-windowed topic analysis; that is, we pick a period of, say,
one month and perform topic analysis within each window.  In addition
to being able to study just which topics were actually worked on and
when, we can also analyze the historical timeline of topics for
patterns and trends.  For example, we might wish to categorize each
topic as being local to a single time window, spanning multiple
consecutive windows, or recurring periodically.  Alternatively, we
might explore evolutionary patterns, such as if topics often recur together, if one particular
topic usually precedes another, if unplanned topics outnumber planned
topics, if particular topics recur near the end of development
iterations, etc. % whereas this paper suggests that most topics are
%local.

@@

% We want to see topics that are unique to a
% certain time period, displayed across the time axis.
% Figure \ref{fig:lda} illustrates the kind of visualization we want to
% automatically generate. 
As part of this work, we wish to explore approaches to automatically
generating visualizations of the topic trends and development timelines, so
that managers may be able to more easily grasp the topic-related
activities of the developers.  Figure \ref{fig:lda} gives an example.
Those topics that
recur, or occur over a larger period are plotted continuously. In our
example we have titled each topic with a word chosen from its
word distribution.  
Given a visualization, such as Figure \ref{fig:lda}, topic
analysis can aid in partitioning a project's time-line into periods of
developer focus.
By breaking apart an iteration into sets of topics
and trends (topics that recur), we may be able to recognize the
underlying software development process and maintenance tasks from these
commits.

@@

Our contributions include:
\begin{itemize}
%\item We demonstrate that windowed topic analysis most topics are local
%  in scope, which would be missed by a global topic analysis.

\item  We introduce windowed topic analysis, and demonstrate its utility
  compared to global topic analysis.

\item We present a number of visualizations of topics and their trends over time to aid
  communication and analysis of these topics and trends.
\item We provide an exploratory case study using these techniques on
  several open source database systems.
\end{itemize}

@@

Lukins et al.~\cite{lukins2008} use LDA to help bug
localization by querying for documents that are related to a bug's
topic. They used LDA to build a topic model and then searched for
similar documents by querying with sample documents.
%These queries would indicate the proportion of
%topics that were associated with the query and the
%related documents.

@@

In addition to finding the topics, LDA can also associate the
documents (i.e., commit messages) in its corpus to one or more of the
topics it discovers.  For example, a commit of a simple bug fix would
probably contain a word such as \emph{ticket}, and thus heavily associate
with the bug fix topic.  A commit that concerned user documentation of
the GUI would likely be associated with both the documentation and GUI
topics if it shared words from both topics.  And a commit that did not
concern bugs, GUIs, or documentation would likely be related to the ``other'' topic.

@@

Once we have our data windowed, we apply our Topic Analysis tool to
each window and extract the top $N$ topics.  We decided to use a
threshold of $N=20$ topics based on our previous experience with LDA; we
found that using more than $20$ makes the boundaries between topics less
distinct.

@@

\shrinkit
\SubSection{Visualization}
\shrinkit
%       1. [ ] need the visualizer
  Visualization is an integral part of our topics analysis which
  allows us to quickly explore the topics and trends of a project.
  Visualizing the results %is desirable as it 
  allows us to explore the
  data from different points of view to address different questions:
  How are the topics spread over time?  How many independent topics
  per period were there?  What were the repeating trends?  Did the
  trends dominate, or did local topics dominate?  How contiguous were
  the trends?  
  To address these questions we will describe and employ multiple visualizations.
%  Since we have a variety of questions to answer we suggest multiple
%  visualizations, which we now describe.

@@

Our choice of the number of topics and adding and removing stop words
produced different results. Our choice of stop words could be biased,
and could affect the results.
%, and our bias for or against such stop words could affect
%the results.
